---
title: "R Markdown for KNIFE"
author: "Haoyuan"
date: "11 March 2019"
output:
  html_document:
    number_sections: yes
    theme: united
    toc: yes
    toc_depth: 3
    toc_float:
      toc_collapsed: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
  word_document:
    toc: yes
    toc_depth: '3'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




```{r results='hide', message=FALSE, warning=FALSE}
library(visNetwork)
library(bnlearn)
library(rbmn)
library(Rgraphviz)
library(gRain)
library(RPostgreSQL)
```

# Data Processing

## Reading data from the database (preprocessed)
```{r results='hide'}
dbdriver <- 'PostgreSQL'
host  <- '127.0.0.1'
port  <- '5432'
user  <- 'postgres'
password <- 'postgres'
dbname <- 'mimic'
schema <- 'mimiciii'

# Connect to the database using the configuration settings
con <- dbConnect(dbDriver(dbdriver), dbname = dbname, host = host, port = port, 
                 user = user, password = password)

# Set the default schema
dbExecute(con, paste("SET search_path TO ", schema, sep=" "))

# Set the pre-processed cohort data
sql_query <- "SELECT * FROM all_data_depression_diabete;"
#data <- dbGetQuery(con, sql_query)
#write.csv(data,file = "/Users/Haoyuan/GoogleDrive/Zhang Haoyuan/KNIFE/MIMIC/data.csv")
```

```{r results='hide'}
data = read.csv("/Users/Haoyuan/GoogleDrive/Zhang Haoyuan/KNIFE/MIMIC/data.csv")
data = data[, -1]
head(data)
```


## Further processing data

```{r results='hide', message=FALSE, warning=FALSE}
tmp0 = data[!duplicated(data[,2]),]
tmp1 = aggregate(hadm_id ~ subject_id, tmp0, function(x) length(unique(x)))
names(tmp1)[2] <- "hadm_count"
tmp2 = aggregate(drug ~ subject_id, tmp0, function(x) length(which(!is.na(x))))
names(tmp2)[2] <- "prescribed"

tmp0 = tmp0[-c(3,4,5,7,8)]
tmp0 = merge(tmp0, tmp1, by ="subject_id")
tmp0 = merge(tmp0, tmp2, by ="subject_id", all = T)
tmp0 = tmp0[-c(2,3)]
tmp0 = tmp0[!duplicated(tmp0[,1]),]
tmp0[is.na(tmp0)]<-0
tmp0$hadm_count = ifelse(tmp0$hadm_count>1, "yes", "no")
names(tmp0)[37] <- "readmission"
tmp0$prescribed = ifelse(tmp0$prescribed==0, "no", "yes")
tmp0$readmission <- as.factor(tmp0$readmission)
tmp0$prescribed <- as.factor(tmp0$prescribed)

```


## Group and discretize variables
```{r}

table(tmp0$ethnicity)
tmp0$ethnicity <- as.character(tmp0$ethnicity)
tmp0$ethnicity[grep("WHITE - BRAZILIAN", tmp0$ethnicity)] = "WHITE"
tmp0$ethnicity[grep("WHITE - EASTERN EUROPEAN", tmp0$ethnicity)] = "WHITE"
tmp0$ethnicity[grep("WHITE - OTHER EUROPEAN", tmp0$ethnicity)] = "WHITE"
tmp0$ethnicity[grep("WHITE - RUSSIAN", tmp0$ethnicity)] = "WHITE"
tmp0$ethnicity[grep("BLACK/AFRICAN AMERICAN", tmp0$ethnicity)] = "BLACK"
tmp0$ethnicity[grep("BLACK/CAPE VERDEAN", tmp0$ethnicity)] = "BLACK"
tmp0$ethnicity[tmp0$ethnicity != "WHITE" & tmp0$ethnicity != "BLACK"]= "OTHERS"
tmp0$ethnicity <- as.factor(tmp0$ethnicity)
table(tmp0$ethnicity)

table(tmp0$marital_status)
tmp0$marital_status <- as.character(tmp0$marital_status)
tmp0$marital_status[tmp0$marital_status != "MARRIED" & tmp0$marital_status != "SINGLE"]= "OTHERS"
tmp0$marital_status <- as.factor(tmp0$marital_status)
table(tmp0$marital_status)


tmp0$age = ifelse(tmp0$age>79, "79+",ifelse(tmp0$age>69,"69_79", ifelse(tmp0$age>59,"59_69", ifelse(tmp0$age>49,"49_59",
                  ifelse(tmp0$age>39,"39_49", ifelse(tmp0$age>29,"29_39", ifelse(tmp0$age>19,"19_29", "19-")))))))
tmp0$age <- as.factor(tmp0$age)
table(tmp0$age)


# tmp0[2] = discretize(tmp0[2], breaks =5, method = "interval")
tmp0$sysbp_max <- as.numeric(tmp0$sysbp_max)
tmp0$diasbp_min<- as.numeric(tmp0$diasbp_min)
tmp0$diasbp_max<- as.numeric(tmp0$diasbp_max)
tmp0$urine_min<- as.numeric(tmp0$urine_min)
tmp0$urine_mean<- as.numeric(tmp0$urine_mean)
tmp0$urine_max<- as.numeric(tmp0$urine_max)
tmp0[,c(7:13, 15:33, 35,36)] = discretize(tmp0[,c(8:13, 15:33, 35,36)], method = 'hartemink', breaks = 4, ibreaks = 5)
tmp0[,c(14, 34)] = discretize(tmp0[,c(14,34)], breaks =3, method = "quantile")

```


## Remove rows with missing data and select a subset of varaibles for exploring purpose
```{r}
model.data = na.omit(tmp0[,c(2:6,16:17,22:23,31,32,34,36, 37:38)])
head(model.data)
```

# Exploratory Data Analysis (EDA) 

feature selection - tbc
```{r}

```



# Building Models
```{r}
# plot network function
# using the visNetwork package to plot the network
plot.network <- function(structure, ht = "400px"){
  nodes.uniq <- unique(c(structure$arcs[,1], structure$arcs[,2]))
  nodes <- data.frame(id = nodes.uniq,
                      label = nodes.uniq,
                      color = "darkturquoise",
                      shadow = TRUE)
  
  edges <- data.frame(from = structure$arcs[,1],
                      to = structure$arcs[,2],
                      arrows = "to",
                      smooth = TRUE,
                      shadow = TRUE,
                      color = "black")
  
  return(visNetwork(nodes, edges, height = ht, width = "100%"))
}


# Structure learning of the BN
dag  = hc(model.data)
plot.network(dag)


#strength of each arcs
strength = arc.strength(dag, model.data, criterion = "x2")
strength
```

In order to further improve the BN strcuture, we can:

## Conditional independence test for each arc

```{r}
ci.test("readmission", "prescribed", "gender", data = model.data, test = "mi")

```

## Add backlist and whitelist before learning the structures - tbc

## Use different structure learning approaches
```{r}
dag1  = hc(model.data)
dag2  = tabu(model.data)
par(mfrow = c(1, 2))
graphviz.compare(dag1, dag2)
```



# Learning Parameters/CPTs

```{r}
est.para <- bn.fit(dag, data = model.data)
est.para
```

```{r fig.width = 3, fig.height = 3}
# Posterior for a single variable
bn.fit.barchart(est.para$readmission, xlab = "Probabilities", ylab = "Levels")

```


```{r fig.width = 10, fig.height = 20}

# Posterior for the network
graphviz.chart(est.para, type = "barprob", grid = TRUE, bar.col = "darkgreen",
               strip.bg = "lightskyblue")
#graphviz.chart(est.para, type = "barprob", grid = TRUE, bar.col = "green", strip.bg = "lightyellow")

```



# Model Validation
```{r warning=FALSE}
#method 1:
#training.set = model.data[1:1487, ]
#test.set = model.data[1488:1859, ]
#bn = naive.bayes(training.set, "readmission")
#fitted = bn.fit(bn, training.set)
#pred = predict(fitted, test.set)
#table(pred, test.set[, "readmission"])

#method 2:
xval = bn.cv(model.data, bn = "hc", loss = "pred-lw",loss.args = list(target = "prescribed"))
xval
OBS = unlist(lapply(xval, `[[`, "observed"))
PRED = unlist(lapply(xval, `[[`, "predicted"))
table(OBS, PRED)

```

compare with other approaches - tbc

# Inference 

## Query for a single variable with evidence
```{r}
cpquery(est.para, event = (readmission=="yes"), evidence = list(prescribed = "yes", age = "59_69"),
        method = "lw")
```

## Forward sampling

    - For a set of varaibles with no evidence
```{r}
cpdist(est.para, nodes = c("readmission", "prescribed", "age", "gender", "marital_status"), evidence = TRUE, n=6)
```

    - For all varaibles with no evidence
```{r}
head(rbn(est.para, 500, model.data))
```

## Rejection sampling with evidence
```{r}
head(cpdist(est.para, nodes = c("readmission", "prescribed", "age"), evidence = (gender == "M")))
```



# Intervention 
```{r}
## ideal interventions and mutilated networks.
dag1 = mutilated(dag, evidence = list(prescribed = "yes"))
plot.network(dag1)

par(mfrow = c(1, 2))
graphviz.compare(dag, dag1)

```

